---
title: "final_project"
author: "Iona Sammons, 300281718"
date: "06/10/2020"
output: html_document
---

# Import stuff
```{r}
library(dplyr)
library(zoo)
library(forecast)
library(ggplot2)
library(lubridate)
library(tseries)

covid <- read.csv("covid_19_data_portal.csv", header = TRUE, stringsAsFactors = FALSE)
```

# Change dataframes
```{r}
# Select data that relates to arrivals
allArrivals <- filter(covid, indicator_name == "Daily border crossings - arrivals")

# Convert values from char to correct format
allArrivals$parameter <- as.Date(allArrivals$parameter, format="%Y-%m-%d")
allArrivals$value <- as.numeric(allArrivals$value)

# Select only other passport arrivals
otherArrivals <- filter(allArrivals, series_name == "Other passports")
otherArrivals <- subset(otherArrivals, select=c("parameter","value"))

####################### converting tests
zoo <- read.zoo(otherArrivals)

tsArrivals<-ts(otherArrivals$value)
plot(tsArrivals)
otherArrivals$cleanvalue<-tsclean(tsArrivals) #remove outliers
```

# Graphs
```{r}
recentArrivals <- filter(otherArrivals, parameter > as.Date("2020-01-01"))

firstReportedCase <- as.Date("2020-02-28")
bordersClosed <- as.Date("2020-03-19")
levelFour <- as.Date("2020-03-25")
MIQStarted <- as.Date("2020-04-10")
levelThree <- as.Date("2020-04-27")
levelTwo <- as.Date("2020-05-13")
levelOne <- as.Date("2020-06-08")
MIQFeeCameIntoEffect <- as.Date("2020-08-11")

ggplot(recentArrivals, aes(x=parameter, y=value)) + 
  geom_line() +
  geom_vline(xintercept = bordersClosed, color = "blue") +
  geom_vline(xintercept = levelFour, color = "red") +
  geom_vline(xintercept = MIQStarted, color = "blue") +
  geom_vline(xintercept = levelThree, color = "dark orange") +
  geom_vline(xintercept = levelTwo, color = "goldenrod1") +
  geom_vline(xintercept = levelOne, color = "green3") +
  geom_vline(xintercept = MIQFeeCameIntoEffect, color = "blue") + 
  labs(title="'Other' passport holders' arrivals since the first reported Covid case",
       #subtitle="",
       y="Number of arrivals",
       x="Month of arrival")
```

```{r}
otherArrivals$Month <- month(otherArrivals$parameter)

ggplot(otherArrivals, aes(parameter, value)) + 
  geom_point() + 
  facet_wrap( ~ Month) + scale_x_date('month')

ggplot(otherArrivals, aes(parameter, cleanvalue)) + 
  geom_point() + 
  facet_wrap( ~ Month) + scale_x_date('month')
```

### Maybe insert col for day of week and do facet wrap for that?

# Moving averages
```{r}
otherArrivals$weeklyma <- ma(otherArrivals$cleanvalue, order=7)
otherArrivals$monthlyma <- ma(otherArrivals$cleanvalue, order=30)

ggplot() + 
  geom_line(otherArrivals, mapping = aes(parameter, cleanvalue, colour="value")) +
  geom_line(otherArrivals, mapping = aes(parameter, weeklyma, colour="weekly ma")) +
  geom_line(otherArrivals, mapping = aes(parameter, monthlyma, colour="monthly ma"))

ggplot() +
  geom_line(otherArrivals, mapping = aes(parameter, weeklyma, colour="weekly ma"))
```

# Seasonal
```{r}
tsArrivals <- ts(otherArrivals$value, frequency = 365)

tsArrivals <- tsclean(tsArrivals)

#remove seasonality/trending
decomp <- stl(tsArrivals, s.window="periodic")

deseasonalArrivals <- seasadj(decomp)

plot(decomp)
```

Augmented Dickey-Fuller. Lag order = 12. Not a big negative number, so I believe that means the data has stationary variance?? Makes sense visually.
```{r}
adf.test(tsArrivals, alternative="stationary")
```

# Correlation
### This section may not be relevent to end report
Acf does not fit within upper and lowerbounds, Pacf does.
```{r}
# NB: 0 = no lag, i.e. the most recent day
Acf(tsArrivals, main="Correlation between data and its lags")
Pacf(tsArrivals, main="Corr between data and its lags that are explained by previous lags")

difference <- diff(deseasonalArrivals, differences = 1)
plot(deseasonalArrivals)
plot(difference)

# ADF confirms if on right track or not
adf.test(difference, alternative = "stationary")

Acf(difference, main="Acf for differenced")
Pacf(difference, main="Pacf for differenced")
```

# Arima models
P=5 D=1 Q=2
```{r}
#fit <- auto.arima(deseasonalArrivals, seasonal=FALSE)
#fit

#tsdisplay(residuals(fit), lag.max=1200)
```



```{r}
tsArrivalsTillDec <- otherArrivalsTillDec <- filter(otherArrivals, parameter<"2019-12-01")
tsArrivalsTillDec <- ts(otherArrivalsTillDec$value, start=2016, frequency = 365)
tsArrivalsTillDec <- tsclean(tsArrivalsTillDec)

tbatsFit <- tbats(tsArrivals, use.parallel=TRUE, num.cores = 2) # fit tbats model
plot(forecast(tbatsFit)) # plot
components <- tbats.components(tbatsFit)
plot(components)

tbatsFit <- tbats(tsArrivalsTillDec, use.parallel=TRUE, num.cores = 2) # fit tbats model
plot(forecast(tbatsFit)) # plot
components <- tbats.components(tbatsFit)
plot(components)
```


# This is confusing me, here's a normal forecast
```{r}
# 305 = 305 days between 01/12/19-30/09/20
plot(forecast(tsArrivalsTillDec, h=305, level=c(80,95)))
```









# Code from original assignment
Added on true false to find out which indicators were not counted
```{r message = F}
library(xts)
# get the time frequency 
# df : subset of covid with one indicator_name value
get.frequency <- function(df) {
  # check if indicator has multiple series
  ind.sub <- df
  if(length(unique(ind.sub$series_name)) > 1 || length(unique(ind.sub$sub_series_name)) > 1){
    ind.sub <- get.single.series(ind.sub)
  }
  # get time frequency of instances with periodicity function
  avg.dif <- periodicity(ind.sub$parameter)
  return(avg.dif$scale)
}
```

The function `get.single.series` removes all but one series / sub series so that only one time period is left.
```{r}
# get the single series for indicator
get.single.series <- function(df) {
  ans <- df %>% filter(series_name == unique(df$series_name)[1])
  # check for multiple sub series
  if(length(unique(ans$sub_series_name)) > 1)
    ans <- ans %>% filter(sub_series_name == unique(df$sub_series_name)[1])
  return(ans)
}
```

Here we create a new data frame of all the indicators and their frequencies. We get a table with frequency counts and `ind.freqs.df`, which can be used to filter the large data set by frequency or lookup the frequency for an indicator.
```{r message = F}
# get instance frequency for all indicators
# convert parameter to date
covid.1 <- covid
covid.1$parameter <- as.Date(covid.1$parameter)
# remove NA dates
covid.1 <- na.omit(covid.1)

indicators <- unique(covid.1$indicator_name)
frequency <- c()
for(i in 1:length(indicators)) {
  ind.sub <- covid.1 %>% filter(indicator_name == indicators[i])
  frequency <- append(frequency, get.frequency(ind.sub))
}
ind.freqs <- cbind(indicators, frequency)
ind.freqs.df <- as.data.frame(ind.freqs)

freq.totals <- ind.freqs.df %>% group_by(frequency) %>% summarise(total = n()) %>% arrange(total)
```
```{r}
library(ggplot2)
ggplot(freq.totals, aes(x=reorder(frequency, total), y=total)) +
  geom_bar(stat="identity", fill = "steelblue") + 
  geom_text(aes(label=total), vjust=1.6, color="white", size=3.5) +
  labs(x = "Frequency", y = "Number of indicators", title = "Instance frequencies of indicators in COVID-19 dataset") +
  theme(panel.background = element_rect(fill = "white", colour = "gray90", size = 1),
          panel.grid.major = element_line(size = 0.7, linetype = 'solid', colour = "gray97"))
```

Find which of the indicators weren't in that graph:
```{r}
indicatorgraphed<- data.frame(ind=unique(covid$indicator_name), contains=unique(covid$indicator_name) %in% ind.freqs)
head(indicatorgraphed)
```

# More code from original assignment
```{r}
# Select data that relates to departures
allArrivals <- filter(covid, indicator_name == "Daily border crossings - arrivals")

# Convert values from char to correct format
allArrivals$parameter <- as.Date(allArrivals$parameter, format="%Y-%m-%d")
allArrivals$value <- as.numeric(allArrivals$value)

# Select NZ passport departures
otherArrivals <- filter(allArrivals, series_name == "Other passports")

# See what we gots...
plot(allArrivals$parameter, allArrivals$value)
plot(otherArrivals$parameter, otherArrivals$value)

# Narrow it down
library(lubridate)
library(stringr)
library(scales)
# Add year only column (needs refactoring):
otherArrivals$year <- year(otherArrivals$parameter)
otherArrivals$year <- as.character(otherArrivals$year)
# Add column with dates altered to all be in 2020 (needs refactoring):
otherArrivals$date_with_uniform_year <- as.character(otherArrivals$parameter)
otherArrivals$date_with_uniform_year <- str_replace(otherArrivals$date_with_uniform_year, "^\\d{4}","2020")
otherArrivals$date_with_uniform_year <- as.Date(otherArrivals$date_with_uniform_year)
# Plot it:
otherArrivals %>%
  ggplot(aes(x = date_with_uniform_year, y = value, colour = year)) +
  geom_line() +
  theme_bw() +
  scale_x_date(labels = date_format("%b")) +
  labs(title="Non-New Zealand passport holders' arrivals over a year",
       subtitle="(Includes both long and short term trips)",
       y="Number of arrivals",
       x="Month of arrival")




firstReportedCase <- as.Date("2020-02-28")
bordersClosed <- as.Date("2020-03-19")
levelFour <- as.Date("2020-03-25")
MIQStarted <- as.Date("2020-04-10")
levelThree <- as.Date("2020-04-27")
levelTwo <- as.Date("2020-05-13")
levelOne <- as.Date("2020-06-08")
MIQFeeCameIntoEffect <- as.Date("2020-08-11")

recentOtherArrivals <- filter(otherArrivals, parameter > firstReportedCase)
ggplot(recentOtherArrivals, aes(x=parameter, y=value)) + 
  geom_line() +
  theme_bw() +
  #geom_vline(xintercept = bordersClosed, color = "blue") +
  geom_vline(xintercept = levelFour, color = "red") +
  #geom_vline(xintercept = MIQStarted, color = "blue") +
  geom_vline(xintercept = levelThree, color = "dark orange") +
  geom_vline(xintercept = levelTwo, color = "goldenrod1") +
  geom_vline(xintercept = levelOne, color = "green3") +
  #geom_vline(xintercept = MIQFeeCameIntoEffect, color = "blue")
  geom_text(mapping = aes(x=levelFour+3, y=10750, label="Level 4"), angle=90) +
  geom_text(mapping = aes(x=levelThree+3, y=10750, label="Level 3"), angle=90) +
  geom_text(mapping = aes(x=levelTwo+3, y=10750, label="Level 2"), angle=90) +
  geom_text(mapping = aes(x=levelOne+3, y=10750, label="Level 1"), angle=90) +
  labs(title="Non-New Zealand passport holders' arrivals since the first reported Covid case",
       #subtitle="",
       y="Number of arrivals",
       x="Month of arrival")
```